
---
title: "Forest Plots (Grayscale): Weighted vs Unweighted"
author: "Generated by M365 Copilot"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    df_print: paged
---

```{r setup, include=FALSE}
# Load libraries
library(readxl)
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
library(glue)

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

## Overview
This version ensures **Weighted** and **Unweighted** ORs and CIs are **clearly separated (no overlap)** and that the `OR [CI]` labels sit **to the right of the bars**. Axis ranges **auto-fit** per plot/panel so nothing is clipped. Race label **Asian → Asian American** is retained.

### Inputs
- Excel file (same folder as this Rmd): `OR_CI_by_group2.xlsx`
- Worksheet: `OR_CI`

```{r read-data}
input_file <- "OR_CI_by_group2.xlsx"
sheet_name <- "OR_CI"

# Read
raw <- readxl::read_excel(input_file, sheet = sheet_name)

# Expected columns
stopifnot(all(c("Model","Race","Variable","OR","CI_Lower","CI_Upper") %in% names(raw)))

# Friendly labels and standardized orders
var_levels <- c("adv_math","non-English only","English only","female","first_gen","frpl","par_exp","par_sch_inv")
var_labels <- c("Advanced math","Non-English only","English only","Female","First-generation","FRPL eligible","Parent expectations","Parent school involvement")

race_levels <- c("Indigenous","Asian American","Black","Latinx","Multiracial","White")

Dat <- raw %>%
  mutate(
    # Weighted/Unweighted flag
    Model_simple = ifelse(grepl("Survey-weighted", Model, fixed = TRUE), "Weighted", "Unweighted"),
    # Rename race: Asian → Asian American
    Race = ifelse(Race == "Asian", "Asian American", Race),
    # Friendly variable labels (fixed order across all plots)
    Variable_label = factor(Variable, levels = var_levels, labels = var_labels),
    # Standardized race order
    Race = factor(Race, levels = race_levels)
  )

# Padding for labels and limits on log scale
pad_point <- 1.03  # label sits just beyond CI upper
pad_limit <- 1.18  # expand x upper limit beyond the max CI for headroom

# Lower bound must be > 0 for log scale
safe_lower <- function(x) max(1e-3, min(x, na.rm = TRUE) / 1.2)
```

## Plotting function (per racial group, auto-fitting range)
```{r plotting-function}
#' Forest plot for a single racial group (grayscale, separated series)
#'
#' Uses vertical dodging by Model to ensure Unweighted vs Weighted do not overlap.
plot_race <- function(race_name, show_labels = TRUE) {
  d <- Dat %>% filter(Race == race_name)
  if (nrow(d) == 0) return(NULL)

  # Per-race axis limits (auto-fit)
  x_lower <- safe_lower(d$CI_Lower)
  x_upper <- max(d$CI_Upper, na.rm = TRUE) * pad_limit

  # Label positions: just past CI upper bound
  d <- d %>% mutate(label_x = pmax(OR, CI_Upper, na.rm = TRUE) * pad_point)

  # Wider dodge to clearly separate series
  dodge_w <- 0.9

  p <- ggplot(d, aes(x = OR, y = Variable_label,
                     color = Model_simple, shape = Model_simple,
                     group = Model_simple)) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
    geom_errorbarh(aes(xmin = CI_Lower, xmax = CI_Upper),
                   position = position_dodge(width = dodge_w),
                   height = 0.12, linewidth = 0.6) +
    geom_point(position = position_dodge(width = dodge_w), size = 2.5, stroke = 0) +
    scale_color_grey(start = 0.25, end = 0.65) +
    scale_shape_manual(values = c("Unweighted" = 16, "Weighted" = 17)) +
    scale_x_log10(
      limits = c(x_lower, x_upper),
      breaks = scales::log_breaks(n = 8),
      labels = label_number(accuracy = 0.01)
    ) +
    coord_cartesian(clip = "off") +
    labs(
      title = glue("Forest plot ({race_name})"),
      x = "Odds ratio (log scale)", y = NULL,
      color = "Model", shape = "Model"
    ) +
    theme_classic(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold",
      panel.border = element_blank(,
      panel.border = element_blank(),
      axis.line = element_line()
    ),
      axis.line = element_line()
    ),
      legend.position = "bottom",
      plot.margin = margin(5.5, 90, 5.5, 5.5),
      panel.border = element_blank(),
      axis.line = element_line()
    )

  if (show_labels) {
    p <- p +
      geom_text(
        aes(x = label_x, label = sprintf("%.2f [%.2f, %.2f]", OR, CI_Lower, CI_Upper),
            group = Model_simple),
        position = position_dodge(width = dodge_w),
        hjust = 0, size = 3, color = "grey10", show.legend = FALSE
      )
  }
  p
}
```

## Forest plots by racial group (auto-fitting, separated series)
```{r plots-by-race, results='asis'}
races <- levels(Dat$Race)
for (r in races) {
  g <- plot_race(r, show_labels = TRUE)
  if (!is.null(g)) print(g)
}
```

## Single multi-panel (faceted) figure (3x2, free x-range, separated series)
```{r faceted-figure}
Dat_for_facets <- Dat %>% mutate(label_x = pmax(OR, CI_Upper, na.rm = TRUE) * pad_point)

dodge_w <- 0.9
p_facets <- ggplot(Dat_for_facets, aes(x = OR, y = Variable_label,
                                      color = Model_simple, shape = Model_simple,
                                      group = Model_simple)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = CI_Lower, xmax = CI_Upper),
                 position = position_dodge(width = dodge_w),
                 height = 0.12, linewidth = 0.5) +
  geom_point(position = position_dodge(width = dodge_w), size = 2.3, stroke = 0) +
  scale_color_grey(start = 0.25, end = 0.65) +
  scale_shape_manual(values = c("Unweighted" = 16, "Weighted" = 17)) +
  scale_x_log10(labels = label_number(accuracy = 0.01)) +
  facet_wrap(~ Race, nrow = 3, ncol = 2, scales = "free_x") +
  coord_cartesian(clip = "off") +
  labs(
    title = "Forest plots by race (grayscale, auto-fit, separated series)",
    subtitle = "Unweighted vs Adjusted (Survey-weighted)",
    x = "Odds ratio (log scale)", y = NULL,
    color = "Model", shape = "Model"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold",
      panel.border = element_blank(),
      axis.line = element_line()
    ),
    legend.position = "bottom",
    plot.margin = margin(5.5, 90, 5.5, 5.5)
  )

# Non-overlapping labels, dodged by model
p_facets <- p_facets +
  geom_text(
    data = Dat_for_facets,
    aes(x = label_x, label = sprintf("%.2f [%.2f, %.2f]", OR, CI_Lower, CI_Upper),
        group = Model_simple),
    position = position_dodge(width = dodge_w),
    hjust = 0, size = 2.8, color = "grey10", show.legend = FALSE
  )

p_facets
```

## (Optional) Save images to PNG
Uncomment to export.

```{r save-each-race, eval=FALSE}
# Save one PNG per racial group
out_dir <- "plots"
if (!dir.exists(out_dir)) dir.create(out_dir)

for (r in levels(Dat$Race)) {
  g <- plot_race(r, show_labels = TRUE)
  if (!is.null(g)) {
    fn <- file.path(out_dir, paste0("forest_", gsub("[\\\\s]", "_", r), "_grayscale.png"))
    ggsave(filename = fn, plot = g, width = 10, height = 6, dpi = 300, bg = "white")
  }
}
```

```{r save-faceted, eval=FALSE}
# Save the multi-panel figure (3 rows x 2 columns, free x scales)
out_dir <- "plots"
if (!dir.exists(out_dir)) dir.create(out_dir)
fn <- file.path(out_dir, "forest_by_race_facet_grayscale_autofit.png")
ggsave(filename = fn, plot = p_facets, width = 12, height = 10, dpi = 300, bg = "white")
```

## Notes
- **Separation:** A wider `position_dodge(width = 0.9)` plus explicit `group = Model_simple` keeps the two series distinct on each row.
- **No overlap with labels:** Labels are placed at `~3%` past the upper CI bound and **dodged** by model, so they don’t collide with bars or each other.
- **Auto-fit axes:** Each plot/panel adjusts x-limits to its data with a small right margin for labels.
