---
title: "analysis"
author: "Kristian Edosomwan"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Prepare for Analysis

## Load libraries
```{r}
library(tidyverse)
library(mice)
library(howManyImputations)
library(gtsummary)
library(survey)
library(jtools)
library(mitools)
```

## Load dataset
```{r}
load("df_imputed.rdata")
df <- haven::read_dta("gcq_hsls_subset.dta")
```

# UNIMPUTED - NO WEIGHTS

### Write function
```{r}
hsls_glm <- function(data) {
  glm(
    p1honors ~ math_4level + 
      par_exp + par_sch_inv +
      female + par_ba + as.factor(first_lang) + frpl +
      math_score + public + as.factor(x1locale),
    data = data,
    family = "quasibinomial"
  )
}


hsls_tbl_merge <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = TRUE, hide_ci = TRUE) %>% 
    add_glance_table() 
}


hsls_tbl_view <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = FALSE, hide_ci = FALSE) %>% 
    add_glance_table() %>% 
    as_flex_table()
}

```
## Asian American
```{r}
asian <- filter(df, stud_race == 1)
asian_glm <- hsls_glm(asian)
asian_glm %>% hsls_tbl_view()
asian_tbl <- asian_glm %>% hsls_tbl_merge
```

## Black
```{r}
# Create Black subset
black <- filter(df, stud_race == 2)

# Run Imputation & check how many needed
black_glm <- black %>% hsls_glm()

# Create Tables
black_tbl <- black_glm %>% hsls_tbl_merge
black_glm %>% hsls_tbl_view()
```

## Indigenous
```{r}
indigenous <- filter(df, stud_race == 3)

# Run Imputation & check how many needed
indigenous_glm <- indigenous %>% hsls_glm()


# Create Tables
indigenous_tbl <- indigenous_glm %>% hsls_tbl_merge
indigenous_glm %>% hsls_tbl_view()
```

## Latinx
```{r}
latinx <- filter(df, stud_race == 4)

# Run Imputation & check how many needed
latinx_glm <- latinx %>% hsls_glm()


# Create Tables
latinx_tbl <- latinx_glm %>% hsls_tbl_merge
latinx_glm %>% hsls_tbl_view()
```

## Two or More Races
```{r}
multi <- filter(df, stud_race == 5)

# Run Imputation & check how many needed
multi_glm <- multi %>% hsls_glm()


# Create Tables
multi_tbl <- multi_glm %>% hsls_tbl_merge
multi_glm %>% hsls_tbl_view()
```

## White
```{r}
white <- filter(df, stud_race == 6)

# Run Imputation & check how many needed
white_glm <- white %>% hsls_glm()


# Create Tables
white_tbl <- white_glm %>% hsls_tbl_merge
white_glm %>% hsls_tbl_view()
```

## Combined Table
```{r}
tbl_merge(
  tbls = list(asian_tbl, black_tbl, indigenous_tbl, latinx_tbl, multi_tbl, white_tbl),
  tab_spanner = c("Asian American", "Black", "Indigenous", "Latinx", "Two or More Races", "White")
  ) %>% 
  as_flex_table()
```

# UNIMPUTED WITH WEIGHTS

## Weight Data
```{r}
library(survey)

df_svy <- svydesign(
  data = df,
  ids = ~stu_id,      # Use the tilde (~) for formulas
  weights = ~w1parent # Use 'weights' instead of 'pweights'
)
df_brr <- svrepdesign(
  weights = ~w1parent, 
  repweights = "w1parent[0-9]+",
  type = "BRR", 
  data = df,
  combined.weights = TRUE
)
```

## Write Function
```{r}
hsls_glm_svy <- function(data) {
  svyglm(
    p1honors ~ math_4level + 
     par_exp + par_sch_inv +
     female + par_ba + factor(first_lang) + frpl +
     math_score + public + factor(x1locale),
    design = data,
    family = "quasibinomial"
  )
}

```
## Asian American
```{r}
asian_svy <- subset(df_brr, stud_race == 1)
asian_svyglm <- hsls_glm_svy(asian_svy)
asian_svyglm %>% hsls_tbl_view()
asian_svytbl <- asian_svyglm %>% hsls_tbl_merge

```

## Black
```{r}
# Create Black subset
black_svy <- subset(df_brr, stud_race == 2)

# Run Imputation & check how many needed
black_svyglm <- black_svy %>% hsls_glm_svy()

# Create Tables
black_svytbl <- black_svyglm %>% hsls_tbl_merge
black_svyglm %>% hsls_tbl_view()
```

## Indigenous
```{r}
indigenous_svy <- subset(df_brr, stud_race == 3)

# Run Imputation & check how many needed
indigenous_svyglm <- indigenous_svy %>% hsls_glm_svy()


# Create Tables
indigenous_svytbl <- indigenous_svyglm %>% hsls_tbl_merge
indigenous_svyglm %>% hsls_tbl_view()
```

## Latinx
```{r}
latinx_svy <- subset(df_brr, stud_race == 4)

# Run Imputation & check how many needed
latinx_svyglm <- latinx_svy %>% hsls_glm_svy()


# Create Tables
latinx_svytbl <- latinx_svyglm %>% hsls_tbl_merge
latinx_svyglm %>% hsls_tbl_view()
```

## Two or More Races
```{r}
multi_svy <- subset(df_brr, stud_race == 5)

# Run Imputation & check how many needed
multi_svyglm <- multi_svy %>% hsls_glm_svy()


# Create Tables
multi_svytbl <- multi_svyglm %>% hsls_tbl_merge
multi_svyglm %>% hsls_tbl_view()
```

## White
```{r}
white_svy <- subset(df_brr, stud_race == 6)

# Run Imputation & check how many needed
white_svyglm <- white_svy %>% hsls_glm_svy()


# Create Tables
white_svytbl <- white_glm %>% hsls_tbl_merge
white_glm %>% hsls_tbl_view()
```

## Combined Table
```{r}
tbl_merge(
  tbls = list(asian_svytbl, black_svytbl, indigenous_svytbl, latinx_svytbl, multi_svytbl, white_svytbl),
  tab_spanner = c("Asian American", "Black", "Indigenous", "Latinx", "Two or More Races", "White")
  ) %>% 
  as_flex_table()
```
## Weighted Analysis for All: Panel Weights
```{r}
glm_all_panel <-       svyglm(design = df_svy,
                        p1honors ~ math_4level + 
                        par_exp + par_race + par_sch_inv +
                        first_gen + first_lang + frpl +
                        math_score + public + urban,
                        family = "quasibinomial")
glm_all_panel %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% as_flex_table()
```

## Weighted Analysis for All: BRR weights
```{r}
glm_all_brr <-       svyglm(design = df_brr,
                        p1honors ~ math_4level + 
                        par_exp + par_race + par_sch_inv +
                        first_gen + first_lang + frpl +
                        math_score + public + urban,
                        family = "quasibinomial")
glm_all_brr %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% as_flex_table()
```
# IMPUTED 

## UNWEIGHTED Racial/Ethnic Subgroups

### Write function
```{r}
hsls_glm_imp <- function(data) {
  with(data,   glm(
    p1honors ~ math_4level + 
      par_exp + par_sch_inv +
      female + par_ba + factor(first_lang) + frpl +
      math_score + public + factor(x1locale),
    family = "quasibinomial"
  ))
}


hsls_tbl_merge <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = TRUE, hide_ci = TRUE) %>% 
    add_glance_table() 
}


hsls_tbl_view <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = FALSE, hide_ci = FALSE) %>% 
    add_glance_table() %>% 
    as_flex_table()
}

```
## Asian American
```{r}
asian_imp <- filter(df_imp, stud_race == 1)
asian_glm <- hsls_glm_imp(asian_imp)
asian_glm %>% hsls_tbl_view()
asian_tbl <- asian_glm %>% hsls_tbl_merge
how_many_imputations(asian_glm)
```

## Black
```{r}
# Create Black subset
black_imp <- filter(df_imp, stud_race == 2)

# Run Imputation & check how many needed
black_glm <- black_imp %>% hsls_glm_imp()
black_glm %>% how_many_imputations()

# Create Tables
black_tbl <- black_glm %>% hsls_tbl_merge
black_glm %>% hsls_tbl_view()
```

## Indigenous
```{r}
indigenous_imp <- filter(df_imp, stud_race == 3)

# Run Imputation & check how many needed
indigenous_glm <- indigenous_imp %>% hsls_glm_imp()
indigenous_glm %>% how_many_imputations()

# Create Tables
indigenous_tbl <- indigenous_glm %>% hsls_tbl_merge
indigenous_glm %>% hsls_tbl_view()
```

## Latinx
```{r}
latinx_imp <- filter(df_imp, stud_race == 4)

# Run Imputation & check how many needed
latinx_glm <- latinx_imp %>% hsls_glm_imp()
latinx_glm %>% how_many_imputations()

# Create Tables
latinx_tbl <- latinx_glm %>% hsls_tbl_merge
latinx_glm %>% hsls_tbl_view()
```

## Two or More Races
```{r}
multi_imp <- filter(df_imp, stud_race == 5)

# Run Imputation & check how many needed
multi_glm <- multi_imp %>% hsls_glm_imp()
multi_glm %>% how_many_imputations()

# Create Tables
multi_tbl <- multi_glm %>% hsls_tbl_merge
multi_glm %>% hsls_tbl_view()
```

## White
```{r}
white_imp <- filter(df_imp, stud_race == 6)

# Run Imputation & check how many needed
white_glm <- white_imp %>% hsls_glm_imp()
white_glm %>% how_many_imputations()

# Create Tables
white_tbl <- white_glm %>% hsls_tbl_merge
white_glm %>% hsls_tbl_view()
```

## Combined Table
```{r}
tbl_merge(
  tbls = list(asian_tbl, black_tbl, indigenous_tbl, latinx_tbl, multi_tbl, white_tbl),
  tab_spanner = c("Asian American", "Black", "Indigenous", "Latinx", "Two or More Races", "White")
  ) %>% 
  as_flex_table()
```

# WEIGHTED Racial/Ethnic Subgroups

## Weight Data

### Create imputation list
```{r}
imp_list <- lapply(df_imp)
```

### Panel Weights
```{r}
library(survey)
library(mitools)

# Creates a list of completed dataset for survey weights
imp_list <- imputationList(
  lapply(1:df_imp$m, function(i) complete(df_imp, i))
)


df_svy_imp <- svydesign(
  data = imp_list,
  ids = ~stu_id,      # Use the tilde (~) for formulas
  weights = ~w1parent # Use 'weights' instead of 'pweights'
)
```

### BRR Weights
```{r}
df_brr_imp <- svrepdesign(
  weights = ~w1parent, 
  repweights = "w1parent[0-9]+",
  type = "BRR", 
  data = imp_list,
  combined.weights = TRUE
)
```

## Write function
```{r}

# function to run model
hsls_glm_svy_imp <- function(data) {
  with(data,   svyglm(
    p1honors ~ as.numeric(math_4level) + 
      as.numeric(par_exp) + as.numeric(par_sch_inv) +
      female + par_ba + factor(first_lang) + frpl +
      math_score + public + factor(x1locale),
    family = "quasibinomial"
  ))
}

# function to exponentiate results into dataframe
make_mi_table <- function(pooled_model) {
  as.data.frame(summary(pooled_model)) %>% 
  tibble::rownames_to_column("variable") %>% # make variable column
  mutate(
    or = exp(results), ll = exp(`(lower`), ul = exp(`upper)`), # exponentiate values
    z = results / se, p_value = 2 * (1 - pnorm(abs(z)))) # calculate p-value with Wald Test
}

```


## Asian American
```{r}
asian_brr_imp <- subset(df_brr_imp, stud_race == 1)

asian_svyglm_imp <- hsls_glm_svy_imp(asian_brr_imp)

asian_svyglm_pooled <- MIcombine(asian_svyglm_imp)

summary(asian_svyglm_pooled)

# extract summary data into table
asian_brr_imp_df <- make_mi_table(asian_svyglm_pooled)

# add race and model
asian_brr_imp_df <- asian_brr_imp_df %>% mutate(race = "Asian American", model = "weighted") # add race and weighted model
```

## Black
```{r}
# Create Black subset
black_brr_imp <- subset(df_brr_imp, stud_race == 2)

black_svyglm_imp <- hsls_glm_svy_imp(black_brr_imp)

black_svyglm_pooled <- MIcombine(black_svyglm_imp)

summary(black_svyglm_pooled)

# extract summary data into table
black_brr_imp_df <- make_mi_table(black_svyglm_pooled)

# add race and model
black_brr_imp_df <- black_brr_imp_df %>% mutate(race = "Black", model = "weighted") # add race and weighted model
```

## Indigenous
```{r}
indigenous_brr_imp <- subset(df_brr_imp, stud_race == 3)

indigenous_svyglm_imp <- hsls_glm_svy_imp(indigenous_brr_imp)

indigenous_svyglm_pooled <- MIcombine(indigenous_svyglm_imp)

summary(indigenous_svyglm_pooled)

# extract summary data into table
indigenous_brr_imp_df <- make_mi_table(indigenous_svyglm_pooled)

# add race and model
indigenous_brr_imp_df <- indigenous_brr_imp_df %>% mutate(race = "indigenous", model = "weighted") # add race and weighted model
```

## Latinx
```{r}
latinx_brr_imp <- subset(df_brr_imp, stud_race == 4)

latinx_svyglm_imp <- hsls_glm_svy_imp(latinx_brr_imp)

latinx_svyglm_pooled <- MIcombine(latinx_svyglm_imp)

summary(latinx_svyglm_pooled)

# extract summary data into table
latinx_brr_imp_df <- make_mi_table(latinx_svyglm_pooled)

# add race and model
latinx_brr_imp_df <- latinx_brr_imp_df %>% mutate(race = "Latinx", model = "weighted") # add race and weighted model
```

## Two or More Races
```{r}
multi_brr_imp <- subset(df_brr_imp, stud_race == 5)

multi_svyglm_imp <- hsls_glm_svy_imp(multi_brr_imp)

multi_svyglm_pooled <- MIcombine(multi_svyglm_imp)

summary(multi_svyglm_pooled)

# extract summary data into table
multi_brr_imp_df <- make_mi_table(multi_svyglm_pooled)

# add race and model
multi_brr_imp_df <- multi_brr_imp_df %>% mutate(race = "Multiracial", model = "weighted") # add race and weighted model
```

## White
```{r}
white_brr_imp <- subset(df_brr_imp, stud_race == 6)

white_svyglm_imp <- hsls_glm_svy_imp(white_brr_imp)

white_svyglm_pooled <- MIcombine(white_svyglm_imp)

summary(white_svyglm_pooled)

# extract summary data into table
white_brr_imp_df <- make_mi_table(white_svyglm_pooled)

# add race and model
white_brr_imp_df <- white_brr_imp_df %>% mutate(race = "White", model = "weighted") # add race and weighted model
```

## Combined Table
```{r}
brr_imp_table <- rbind(asian_brr_imp_df, black_brr_imp_df)
brr_imp_table <- rbind(brr_imp_table, indigenous_brr_imp_df)
brr_imp_table <- rbind(brr_imp_table, latinx_brr_imp_df)
brr_imp_table <- rbind(brr_imp_table, multi_brr_imp_df)
brr_imp_table <- rbind(brr_imp_table, white_brr_imp_df)
write_csv(brr_imp_table, "weighted_imputed_table_results.csv")

```