---
title: "Prepare HSLS GCQ Data for Analysis"
author: "Kristian Edosomwan"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Data Engineering
## Load Libraries and Datasets

The full raw dataset (n = 23,503)
```{r libraries}
library(haven)
library(dplyr)
library(labelled)
library(psych)
library(broom)
library(gtsummary)
library(jtools)
hsls_raw <- read_dta("hsls_17_student_pets_sr_v1_0.dta")
names(hsls_raw) <- tolower(names(hsls_raw))
```

## Create Analytic Sample
after removing parent non-responders (n = 16788)
after removing no math students (n = 14805)
```{r remove non-response}
hsls_raw %>% count(p1honors)
hsls_raw %>% count(x3thimath9)

# take out observations with outcome non-response
hsls_na <- hsls_raw %>% filter(p1honors !=-8)
# take out observations where student is not in a mathematics class
hsls_na <- hsls_na %>% filter(s1mfall09 == 1) 
```

## summarize original variables in model
```{r}
hsls_model <- hsls_na %>% dplyr::select(p1honors, x3thimath, x1poverty185, x1duallang, x1control, x1locale, x1paredu, x1sex, x1race, p1ptomtg, p1schmtg, p1schevent, p1volunteer, p1fundraise, p1ptconfer, ends_with("m09"), x1txmtscor, x1txmquint)

hsls_model %>% drop_unused_value_labels() %>% haven::as_factor() %>%  tbl_summary()
```

```{r subset replace missing values with NA}
# select relevant variables
hsls_na <- hsls_na %>% select(stu_id, starts_with("x1"), starts_with("p1"), x3thimath9, ends_with("m09"), starts_with("w1par"))

# replace missing values
hsls_na[hsls_na == -9] <- NA
hsls_na[hsls_na == -8] <- NA
hsls_na[hsls_na == -7] <- NA
hsls_na[hsls_na == -6] <- NA
hsls_na[hsls_na == -5] <- NA

# filter out variables where more than 10% is missing
hsls_na <- hsls_na %>% dplyr::select(where(~mean(is.na(.)) <= 0.10))
```

## Variable Manipulation

Decided to use student survey to determine math level because it has less missing data than the transcript data for 9th grade math class. Using Math course indicators based on Riegle-Crumb et al. (2005, p. 14)
```{r}
# ---------------------------
# Treatment variables
# ---------------------------
hsls <- hsls_na %>%
  mutate(
      math9 = ordered(
        case_when(
      s1alg1m09  == 1 ~ "Algebra I",
      s1geom09   == 1 ~ "Geometry",
      s1alg2m09  == 1 ~ "Algebra II",
      s1trigm09  == 1 ~ "Trigonometry",
      s1revm09   == 1 ~ "Remedial Math",
      s1statsm09 == 1 ~ "Statistics/Probability",
      s1prealgm09== 1 ~ "Pre-Algebra",
      s1angeom09 == 1 ~ "Analytic Geometry",
      s1advm09   == 1 ~ "Advanced Math",
      s1othm09   == 1 ~ "Other Math",
      TRUE ~ NA_character_
    ),
    levels = c("Other Math","Remedial Math", "Pre-Algebra", "Algebra I", "Geometry", "Analytic Geometry", "Algebra II", "Advanced Math","Statistics/Probability", "Trigonometry")),

    # adv_math: binary numeric (0 = not advanced, 1 = advanced)
    adv_math = ifelse(math9 >=5, 1, 0),
    # math_3level: ordinal categorical (low → high)
    math_3level = ordered(
      case_when(
        math9 <= 3 ~ "remedial",
        math9 == 4 ~ "on-level",
        math9 >= 5 ~ "advanced"
      ),
      levels = c("remedial", "on-level", "advanced")
    ),

    # math_4level: ordinal categorical (low → high)
    math_4level = ordered(
      case_when(
        as.numeric(math9) <= 3 ~ "remedial",
        as.numeric(math9) == 4 ~ "on-level",
        as.numeric(math9) == 5 | as.numeric(math9) == 6 ~ "advanced",
        as.numeric(math9) >= 7 ~ "double advanced"
      ),
      levels = c("remedial", "on-level", 
                 "advanced", "double advanced")
    )
  )

# ---------------------------
# Controls
# ---------------------------
hsls <- hsls %>%
  mutate(
    # urban: binary numeric
    urban = ifelse(x1locale == 1, 1, 0),

    # math_score: continuous numeric
    math_score = ifelse(x1txmquint < 0, NA, x1txmtscor),
    math_quintile = x1txmquint,

    # public: binary numeric (1 = public, 0 = private)
    public = ifelse(x1control == 1, 1, 0)
  )

# ---------------------------
# Demographics
# ---------------------------
hsls <- hsls %>%
  mutate(
    # first_gen: binary numeric
    first_gen = ifelse(x1paredu <= 3,1,0),
    par_ba = ifelse(x1paredu >= 4, 1, 0),

    # female: binary numeric (1 = female)
    female = ifelse(x1sex == 2, 1, 0),

    # frpl: binary numeric
    frpl = ifelse(x1poverty185 == 0, 0, 1),

    # pov_level: ordinal numeric (1 = deepest poverty → 4 = no poverty)
    pov_level = ordered(
      case_when(
        x1poverty185 == 0  ~ 4,
        x1poverty185 == 1 & x1poverty130 == 0 ~ 3,
        x1poverty130 == 1 & x1poverty == 0 ~ 2,
        x1poverty == 1 ~ 1
      ),
      levels = c(1, 2, 3, 4)
    ),

    # first_lang: nominal categorical
    first_lang = factor(
      case_when(
        x1duallang == 1 ~ "English only",
        x1duallang == 2 ~ "Non-English only",
        x1duallang == 3 ~ "English and another equally"
      )
    ),

    # esl_program: binary numeric
    esl_program = case_when(
      p1ellever == 1 ~ "1 yes",
      p1ellever == 2 ~ "0 no"
    ),

    # stud_race: nominal categorical
    stud_race = factor(
      case_when(
        x1race %in% c(1, 7) ~ "Indigenous",
        x1race == 2        ~ "Asian",
        x1race == 3        ~ "Black",
        x1race %in% c(4,5) ~ "Latinx",
        x1race == 6        ~ "Multiple",
        x1race == 8        ~ "White"
      )
    ),

    # par_race: nominal categorical
    par_race = factor(
      case_when(
        x1par1race == 2        ~ "Asian",
        x1par1race == 3        ~ "Black",
        x1par1race %in% c(4,5) ~ "Latinx",
        x1par1race == 6        ~ "Multiple",
        x1par1race == 8        ~ "White",
        x1par1race == 9        ~ "Indigenous"
      )
    )
  )

# ---------------------------
# Parent involvement
# ---------------------------
hsls <- hsls %>%
  mutate(

    # par_sch_inv: count variable (numeric, Poisson-type)
    par_sch_inv = as.numeric(p1ptomtg +
                  p1schmtg +
                  p1schevent +
                  p1volunteer +
                  p1fundraise +
                  p1ptconfer)
  )

# ---------------------------
# Parent expectations
# ---------------------------
hsls <- hsls %>%
  mutate(
    exp_ba = ifelse(p1eduexpect >= 6, 1, 0),
    # par_exp: ordinal numeric (low → high expectations)
    par_exp = ordered(
      case_when(
        p1eduexpect == 11            ~ 1, # IDK
        p1eduexpect %in% c(1,2,3,4)  ~ 2, # high school to Associate
        p1eduexpect %in% c(5,6)      ~ 3, # start/complete BA
        p1eduexpect %in% c(7,8)      ~ 4, #start/complete MA
        p1eduexpect %in% c(9,10)     ~ 5 # sart/complete phd/professional
      ),
      levels = c(1, 2, 3, 4, 5)
    )
  )

# check poverty groups
##hsls %>% dplyr::select(pov_level, starts_with("x1pov")) %>% tbl_summary(by = pov_level)
```

## Descriptive Statistics

```{r}
df_summ <- hsls %>% dplyr::select(p1honors, math_4level,
                        par_exp, par_ba, par_sch_inv,
                        first_lang, frpl, stud_race,
                        math_score, public, x1locale)

df_summ %>% gtsummary::tbl_summary(by = stud_race, type = list(all_continuous() ~ "continuous2",
                          c(par_sch_inv) ~ "continuous2"),
    statistic = all_continuous() ~ "{mean} ({sd})") %>% as_flex_table()
```
## Save Clean Dataset

```{r save}
write_dta(hsls, "gcq_hsls_subset.dta")
```

# Analysis & Model Building on Unimputed Data

## Weight Data
```{r}
library(survey)

df_svy <- svydesign(
  data = hsls,
  ids = ~stu_id,      # Use the tilde (~) for formulas
  weights = ~w1parent # Use 'weights' instead of 'pweights'
)
df_brr <- svrepdesign(
  weights = ~w1parent, 
  repweights = "w1parent[0-9]+",
  type = "BRR", 
  data = hsls,
  combined.weights = TRUE
)
```

## Run Data for All 3 Math Levels
```{r}
glm_all3 <-glm(data = hsls,
           p1honors ~ math_3level + 
           par_exp + par_race + par_sch_inv +
           first_gen + first_lang + frpl +
           math_score + public + urban,
           family = "binomial")
jtools::summ(glm_all3, exp = TRUE)
glm_all3 %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% add_glance_table() %>% as_flex_table()
```

## Run Data for All 4 Math Levels
```{r}
glm_all4 <-glm(data = hsls,
           p1honors ~ math_4level + 
           par_exp + par_race + par_sch_inv +
           first_gen + first_lang + frpl +
           math_score + public + urban,
           family = "binomial")
jtools::summ(glm_all4, exp = TRUE)


glm_all4 %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% add_glance_table() %>% as_flex_table()
```

## Run Data for All 4 Math Levels with quintile instead of score
```{r}
glm_all4_quint <-glm(data = hsls,
           p1honors ~ math_4level + 
           par_exp + par_race + par_sch_inv +
           first_gen + first_lang + frpl +
           ordered(math_quintile) + public + urban,
           family = "binomial")
jtools::summ(glm_all4_quint, exp = TRUE)


glm_all4_quint %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% add_glance_table() %>% as_flex_table()
```

## Weighted Analysis for All: Panel Weights
```{r}
glm_all_panel <-       svyglm(design = df_svy,
                        p1honors ~ math_3level + 
                        par_exp + par_race + par_sch_inv +
                        first_gen + first_lang + frpl +
                        math_score + public + urban,
                        family = "binomial")
glm_all_panel %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% as_flex_table()
```

## Weighted Analysis for All: BRR weights
```{r}
glm_all_brr <-       svyglm(design = df_brr,
                        p1honors ~ math_3level + 
                        par_exp + par_race + par_sch_inv +
                        first_gen + first_lang + frpl +
                        math_score + public + urban,
                        family = "binomial")
glm_all_brr %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% as_flex_table()
```





