---
title: "analysis"
author: "Kristian Edosomwan"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Prepare for Analysis

## Load libraries
```{r}
library(tidyverse)
library(mice)
library(howManyImputations)
library(gtsummary)
library(survey)
library(jtools)
library(mitools)
```

## Load datasets
```{r}
load("df_imputed.rdata")
df <- haven::read_dta("gcq_hsls_subset.dta")
```
# Descriptive Statistics

```{r}
df_summ <- df %>% labelled::drop_unused_value_labels() %>%  haven::as_factor() %>% dplyr::select(p1honors, math_4level,
                        par_exp, female, par_ba, par_sch_inv,
                        first_lang, frpl, stud_race,
                        math_score, public, x1locale)

df_summ %>% gtsummary::tbl_summary(by = stud_race, type = list(all_continuous() ~ "continuous2",
                          c(par_sch_inv) ~ "continuous2"),
    statistic = all_continuous() ~ "{mean} ({sd})") %>% as_flex_table()
```
# UNIMPUTED

## NO WEIGHTS
### Write function
```{r}
hsls_glm <- function(data) {
  glm(
    p1honors ~ as.numeric(math_4level) + 
      as.numeric(par_exp) + as.numeric(par_sch_inv) +
      female + par_ba + as.factor(first_lang) + frpl +
      math_score + public + as.factor(x1locale),
    data = data,
    family = "binomial"
  )
}


hsls_tbl_merge <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = TRUE, hide_ci = TRUE) %>% 
    add_glance_table() 
}


hsls_tbl_view <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = FALSE, hide_ci = FALSE) %>% 
    add_glance_table() %>% 
    as_flex_table()
}

```
### Asian American
```{r}
asian <- filter(df, stud_race == 1)
asian_glm <- hsls_glm(asian)
asian_glm %>% hsls_tbl_view()
asian_tbl <- asian_glm %>% hsls_tbl_merge
```

### Black
```{r}
# Create Black subset
black <- filter(df, stud_race == 2)

# Run Imputation & check how many needed
black_glm <- black %>% hsls_glm()

# Create Tables
black_tbl <- black_glm %>% hsls_tbl_merge
black_glm %>% hsls_tbl_view()
```

### Indigenous
```{r}
indigenous <- filter(df, stud_race == 3)

# Run Imputation & check how many needed
indigenous_glm <- indigenous %>% hsls_glm()


# Create Tables
indigenous_tbl <- indigenous_glm %>% hsls_tbl_merge
indigenous_glm %>% hsls_tbl_view()
```

### Latinx
```{r}
latinx <- filter(df, stud_race == 4)

# Run Imputation & check how many needed
latinx_glm <- latinx %>% hsls_glm()


# Create Tables
latinx_tbl <- latinx_glm %>% hsls_tbl_merge
latinx_glm %>% hsls_tbl_view()
```

### Two or More Races
```{r}
multi <- filter(df, stud_race == 5)

# Run Imputation & check how many needed
multi_glm <- multi %>% hsls_glm()


# Create Tables
multi_tbl <- multi_glm %>% hsls_tbl_merge
multi_glm %>% hsls_tbl_view()
```

### White
```{r}
white <- filter(df, stud_race == 6)

# Run Imputation & check how many needed
white_glm <- white %>% hsls_glm()


# Create Tables
white_tbl <- white_glm %>% hsls_tbl_merge
white_glm %>% hsls_tbl_view()
```


### Combined Tables
```{r}
# combined table publication ready with a few modifications
tbl_merge(
  tbls = list(asian_tbl, black_tbl, indigenous_tbl, latinx_tbl, multi_tbl, white_tbl),
  tab_spanner = c("Asian American", "Black", "Indigenous", "Latinx", "Two or More Races", "White")
  ) %>% 
  as_flex_table()

# combined table with model fit 
export_summs(
  asian_glm, black_glm, indigenous_glm, latinx_glm, multi_glm, white_glm, 
  model.names = c("Asian American", "Black", "Indigenous", "Latinx", "Two or More Races", "White"), 
  statistics = "all"
  )
```


## WEIGHTS
### Weight Data
```{r}
library(survey)

df_brr <- svrepdesign(
  weights = ~w1parent, 
  repweights = "w1parent[0-9]+",
  type = "BRR", 
  data = df,
  combined.weights = TRUE
)


```

### Write Function
```{r}
hsls_glm_svy <- function(data) {
  svyglm(
    p1honors ~ as.numeric(math_4level) + 
      as.numeric(par_exp) + as.numeric(par_sch_inv) +
      female + par_ba + as.factor(first_lang) + frpl +
      math_score + public + as.factor(x1locale),
      design = data,
    family = "quasibinomial"
  )
}

```
### Asian American
```{r}
asian_svy <- subset(df_brr, stud_race == 1)

asian_svyglm <- hsls_glm_svy(asian_svy)
asian_svyglm %>% poliscidata::fit.svyglm()


asian_svyglm %>% hsls_tbl_view()
asian_svytbl <- asian_svyglm %>% hsls_tbl_merge

```

### Black
```{r}
# Create Black subset
black_svy <- subset(df_brr, stud_race == 2)

# Run model and get R-squared
black_svyglm <- black_svy %>% hsls_glm_svy()
black_svyglm %>% poliscidata::fit.svyglm()

# Create Tables
black_svytbl <- black_svyglm %>% hsls_tbl_merge
black_svyglm %>% hsls_tbl_view()
```

### Indigenous - Doesn't converge so not included in document
```{r, eval = FALSE}
indigenous_svy <- subset(df_brr, stud_race == 3)

# Run model and get R-squared
indigenous_svyglm <- indigenous_svy %>% hsls_glm_svy()


# Create Tables
indigenous_svytbl <- indigenous_svyglm %>% hsls_tbl_merge
indigenous_svyglm %>% hsls_tbl_view()
```

### Latinx
```{r}
latinx_svy <- subset(df_brr, stud_race == 4)

# Run Imputation & check how many needed
latinx_svyglm <- latinx_svy %>% hsls_glm_svy()
latinx_svyglm %>% poliscidata::fit.svyglm()

# Create Tables
latinx_svytbl <- latinx_svyglm %>% hsls_tbl_merge
latinx_svyglm %>% hsls_tbl_view()
```

### Two or More Races
```{r}
multi_svy <- subset(df_brr, stud_race == 5)

# Run Imputation & check how many needed
multi_svyglm <- multi_svy %>% hsls_glm_svy()
multi_svyglm %>% poliscidata::fit.svyglm()

# Create Tables
multi_svytbl <- multi_svyglm %>% hsls_tbl_merge
multi_svyglm %>% hsls_tbl_view()
```

### White
```{r}
white_svy <- subset(df_brr, stud_race == 6)

# Run Imputation & check how many needed
white_svyglm <- white_svy %>% hsls_glm_svy()
white_svyglm %>% poliscidata::fit.svyglm()

# Create Tables
white_svytbl <- white_svyglm %>% hsls_tbl_merge
white_glm %>% hsls_tbl_view()
```

### Combined Table
```{r}

# publication ready table
tbl_merge(
  tbls = list(asian_svytbl, black_svytbl, latinx_svytbl, multi_svytbl, white_svytbl),
  tab_spanner = c("Asian American", "Black", "Latinx", "Two or More Races", "White")
  ) %>% 
  as_flex_table()


```

# IMPUTED 

## UNWEIGHTED Racial/Ethnic Subgroups

### Write function
```{r}
hsls_glm_imp <- function(data) {
  with(data,     
       glm(
         p1honors ~ as.numeric(math_4level) + 
         as.numeric(par_exp) + as.numeric(par_sch_inv) +
         female + par_ba + as.factor(first_lang) + frpl +
         math_score + public + as.factor(x1locale),
         family = "binomial"
  ))
}

hsls_tbl_imp_merge <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = TRUE, hide_ci = TRUE) %>% 
    add_glance_table() 
}


hsls_tbl_imp_view <- function(data){
  data %>% 
    tbl_regression(exp = TRUE) %>% 
    add_significance_stars(hide_se = FALSE, hide_p = FALSE, hide_ci = FALSE) %>% 
    add_glance_table() %>% 
    as_flex_table()
}

make_imp_table <- function(model) {
  as.data.frame(summary(pool(model))) %>% # make data frame or pooled model summary
    tibble::rownames_to_column("variable") %>% # make variable column
    dplyr::mutate(
      or = exp(estimate), # exponentiate values for ORs
      ll = exp(estimate - (qnorm(0.975) * std.error)), # exponentiate lower limit with Wald interval
      ul = exp(estimate + (qnorm(0.975) * std.error)), # exponentiate upper limit with Wald interval
      se = std.error) %>% 
    dplyr::select(or, se, ll, ul, p.value)  
}

```
### Asian American
```{r}
# create subset
asian_imp <- filter(df_imp, stud_race == 1)

# Run Imputated model & check how many needed
asian_glm_imp <- asian_imp %>% hsls_glm_imp()
how_many_imputations(asian_glm_imp)

# Create Tables
asian_imp_df <- asian_glm_imp %>%  make_imp_table() %>% mutate(race = "Asian American", model = "unweighted") # add race and weighted model
asian_tbl_imp <- asian_glm_imp %>% hsls_tbl_imp_merge
asian_glm_imp %>% hsls_tbl_imp_view()
```

### Black
```{r}
# Create Black subset
black_imp <- filter(df_imp, stud_race == 2)

# Run Imputation & check how many needed
black_glm_imp <- black_imp %>% hsls_glm_imp()
black_glm_imp %>% how_many_imputations()

# Create Tables
black_imp_df <- black_glm_imp %>%  make_imp_table()  %>% mutate(race = "Black", model = "unweighted") # add race and weighted model
black_tbl_imp <- black_glm_imp %>% hsls_tbl_imp_merge
black_glm_imp %>% hsls_tbl_imp_view()
```

### Indigenous
```{r}
indigenous_imp <- filter(df_imp, stud_race == 3)

# Run Imputation & check how many needed
indigenous_glm_imp <- indigenous_imp %>% hsls_glm_imp()
indigenous_glm_imp %>% how_many_imputations()

# Create Tables
indigenous_imp_df <- indigenous_glm_imp %>%  make_imp_table()  %>% mutate(race = "indigenous", model = "unweighted") # add race and weighted model
indigenous_tbl_imp <- indigenous_glm_imp %>% hsls_tbl_imp_merge
indigenous_glm_imp %>% hsls_tbl_imp_view()
```

### Latinx
```{r}
latinx_imp <- filter(df_imp, stud_race == 4)

# Run Imputation & check how many needed
latinx_glm_imp <- latinx_imp %>% hsls_glm_imp()
latinx_glm_imp %>% how_many_imputations()

# Create Tables
latinx_imp_df <- latinx_glm_imp %>%  make_imp_table()  %>% mutate(race = "latinx", model = "unweighted") # add race and weighted model
latinx_tbl_imp <- latinx_glm_imp %>% hsls_tbl_imp_merge
latinx_glm_imp %>% hsls_tbl_imp_view()
```

### Two or More Races
```{r}
multi_imp <- filter(df_imp, stud_race == 5)

# Run Imputation & check how many needed
multi_glm_imp <- multi_imp %>% hsls_glm_imp()
multi_glm_imp %>% how_many_imputations()

# Create Tables
multi_imp_df <- multi_glm_imp %>%  make_imp_table()  %>% mutate(race = "Multiracial", model = "unweighted") # add race and weighted model
multi_tbl_imp <- multi_glm_imp %>% hsls_tbl_imp_merge
multi_glm_imp %>% hsls_tbl_imp_view()
```

### White
```{r}
white_imp <- filter(df_imp, stud_race == 6)

# Run Imputation & check how many needed
white_glm_imp <- white_imp %>% hsls_glm_imp()
white_glm_imp %>% how_many_imputations()

# Create Tables
white_imp_df <- white_glm_imp %>%  make_imp_table() %>% mutate(race = "White", model = "unweighted") # add race and weighted model
white_tbl_imp <- white_glm_imp %>% hsls_tbl_imp_merge
white_glm_imp %>% hsls_tbl_imp_view()
```

### Combined Table
```{r}
tbl_merge(
  tbls = list(asian_tbl_imp, black_tbl_imp, indigenous_tbl_imp, latinx_tbl_imp, multi_tbl_imp, white_tbl_imp),
  tab_spanner = c("Asian American", "Black", "Indigenous", "Latinx", "Two or More Races", "White")
  ) %>% 
  as_flex_table()

#csv
imp_table <- rbind(asian_imp_df, black_imp_df)
imp_table <- rbind(imp_table, indigenous_imp_df) # commenting out since indigenous model did not converge
imp_table <- rbind(imp_table,latinx_imp_df)
imp_table <- rbind(imp_table, multi_imp_df)
imp_table <- rbind(imp_table, white_imp_df)
write_csv(imp_table, "unweighted_imputed_table_results.csv")
```


## WEIGHTED Racial/Ethnic Subgroups
### BRR Weight Data
```{r}
library(survey)
library(mitools)

# Creates a list of completed dataset for survey weights
imp_list <- imputationList(
  lapply(1:df_imp$m, function(i) complete(df_imp, i))
)

# creates BRR survey design
df_brr_imp <- svrepdesign(
  weights = ~w1parent, 
  repweights = "w1parent[0-9]+",
  type = "BRR", 
  data = imp_list,
  combined.weights = TRUE
) 
```

### Write function
```{r}

# function to run model
hsls_glm_svy_imp <- function(data) {
  with(data,   
       svyglm(
       p1honors ~ as.numeric(math_4level) + 
       as.numeric(par_exp) + as.numeric(par_sch_inv) +
       female + par_ba + as.factor(first_lang) + frpl +
       math_score + public + as.factor(x1locale),
       family = "quasibinomial"
  ))
}

# function to exponentiate results and create dataframe with or, se, ll, ul, & p-value
make_mi_table <- function(model) {
  as.data.frame(summary(MIcombine(model))) %>%  # make dataframe of pooled model
  tibble::rownames_to_column("variable") %>% # make variable column
  mutate(
    or = exp(results), # exponentiate estimate for ORs
    ll = exp(`(lower`), # exponentiate ll
    ul = exp(`upper)`), # exponentiate ul
    z = results / se, 
    p.value = 2 * (1 - pnorm(abs(z))) %>%  # calculate p-value with Wald Test
  dplyr::select(or, se, ll, ul, p.value)
  )
}

```


### Asian American
```{r}
# subset data
asian_brr_imp <- subset(df_brr_imp, stud_race == 1)

# run weighted imputed data
asian_svyglm_imp <- hsls_glm_svy_imp(asian_brr_imp)

# extract summary data into table
asian_brr_imp_df <- make_mi_table(asian_svyglm_imp)
asian_brr_imp_df

# add race and model
asian_brr_imp_df <- asian_brr_imp_df %>% mutate(race = "Asian American", model = "weighted") # add race and weighted model
```

### Black
```{r}
# Create Black subset
black_brr_imp <- subset(df_brr_imp, stud_race == 2)

# run weighted imputed model
black_svyglm_imp <- hsls_glm_svy_imp(black_brr_imp)

# extract summary data into table
black_brr_imp_df <- make_mi_table(black_svyglm_imp)
black_brr_imp_df

# add race and model
black_brr_imp_df <- black_brr_imp_df %>% mutate(race = "Black", model = "weighted") # add race and weighted model
```

### Indigenous - Does not converge so not evaluated in document
```{r, eval = FALSE}
indigenous_brr_imp <- subset(df_brr_imp, stud_race == 3)

indigenous_svyglm_imp <- hsls_glm_svy_imp(indigenous_brr_imp)

# extract summary data into table
indigenous_brr_imp_df <- make_mi_table(indigenous_svyglm_imp)

# add race and model
indigenous_brr_imp_df <- indigenous_brr_imp_df %>% mutate(race = "indigenous", model = "weighted") # add race and weighted model
```

### Latinx
```{r}
latinx_brr_imp <- subset(df_brr_imp, stud_race == 4)

latinx_svyglm_imp <- hsls_glm_svy_imp(latinx_brr_imp)

# extract summary data into table
latinx_brr_imp_df <- make_mi_table(latinx_svyglm_imp)
latinx_brr_imp_df

# add race and model
latinx_brr_imp_df <- latinx_brr_imp_df %>% mutate(race = "Latinx", model = "weighted") # add race and weighted model
```

### Two or More Races
```{r}
multi_brr_imp <- subset(df_brr_imp, stud_race == 5)

multi_svyglm_imp <- hsls_glm_svy_imp(multi_brr_imp)

# extract summary data into table
multi_brr_imp_df <- make_mi_table(multi_svyglm_imp)
multi_brr_imp_df

# add race and model
multi_brr_imp_df <- multi_brr_imp_df %>% mutate(race = "Multiracial", model = "weighted") # add race and weighted model
```

### White
```{r}
white_brr_imp <- subset(df_brr_imp, stud_race == 6)

white_svyglm_imp <- hsls_glm_svy_imp(white_brr_imp)

# extract summary data into table
white_brr_imp_df <- make_mi_table(white_svyglm_imp)
white_brr_imp_df

# add race and model
white_brr_imp_df <- white_brr_imp_df %>% mutate(race = "White", model = "weighted") # add race and weighted model
```

### Combined Table
```{r}
brr_imp_table <- rbind(asian_brr_imp_df, black_brr_imp_df)
# brr_imp_table <- rbind(brr_imp_table, indigenous_brr_imp_df) # commenting out since indigenous model did not converge
brr_imp_table <- rbind(brr_imp_table, latinx_brr_imp_df)
brr_imp_table <- rbind(brr_imp_table, multi_brr_imp_df)
brr_imp_table <- rbind(brr_imp_table, white_brr_imp_df)
write_csv(brr_imp_table, "weighted_imputed_table_results.csv")

# combine weighted and unweighted tables
all_imp_table <- rbind(brr_imp_table, imp_table)
write_csv(all_imp_table, "all_imputed_table_results.csv")
```