---
title: "Prepare HSLS GCQ Data for Analysis"
author: "Kristian Edosomwan"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Data Engineering
## Load Libraries and Datasets

The full raw dataset (n = 23,503)
```{r libraries}
library(haven)
library(dplyr)
library(labelled)
library(psych)
library(broom)
library(gtsummary)
library(jtools)
hsls_raw <- read_dta("hsls_17_student_pets_sr_v1_0.dta")
names(hsls_raw) <- tolower(names(hsls_raw))
```

## Create Analytic Sample
after removing parent non-responders (n = 16788)
after removing no math students (n = 14805)
```{r remove non-response}
hsls_raw %>% count(p1honors)
hsls_raw %>% count(x3thimath9)

# take out observations with outcome non-response
hsls_na <- hsls_raw %>% filter(p1honors !=-8)
# take out observations where student is not in a mathematics class
hsls_na <- hsls_na %>% filter(s1mfall09 == 1) 
```

## summarize original variables in model
```{r}
hsls_model <- hsls_na %>% dplyr::select(p1honors, x3thimath, x1poverty185, x1duallang, x1control, x1locale, x1paredu, x1sex, x1race, p1ptomtg, p1schmtg, p1counselor, p1schevent, p1volunteer, p1fundraise, p1ptconfer, ends_with("m09"), x1txmtscor, x1txmquint)

hsls_math <- hsls_raw %>% dplyr::select(ends_with("m09"))

hsls_par <- hsls_raw %>% dplyr::select(p1honors, p1ptomtg, p1schmtg, p1counselor, p1schevent, p1volunteer, p1fundraise, p1ptconfer)

## running EFA on the parent related variables
# hsls_par[hsls_par == -9] <- NA
# hsls_par[hsls_par == -8] <- NA
# hsls_par[hsls_par == -7] <- NA
# hsls_par[hsls_par == -6] <- NA
# hsls_par[hsls_par == -5] <- NA
# EFAtools::N_FACTORS(hsls_par)
# EFAtools::EFA(hsls_par, n_factors = 2)
```

```{r subset replace missing values with NA}
# select relevant variables
hsls_na <- hsls_na %>% select(stu_id, starts_with("x1"), starts_with("p1"), x3thimath9, starts_with("w1par"))

# replace missing values
hsls_na[hsls_na == -9] <- NA
hsls_na[hsls_na == -8] <- NA
hsls_na[hsls_na == -7] <- NA
hsls_na[hsls_na == -6] <- NA
hsls_na[hsls_na == -5] <- NA

# filter out variables where more than 10% is missing
hsls_na <- hsls_na %>% dplyr::select(where(~mean(is.na(.)) <= 0.10))
```

## Variable Manipulation
```{r}
# ---------------------------
# Outcome variable
# ---------------------------
# par_says_honors: binary numeric (0/1)
hsls <- hsls_na %>%
  mutate(
    par_says_honors = case_when(
      p1honors == 1 ~ "1 yes",
      p1honors == 0 ~ "0 no",
      TRUE ~ NA
    )
  )

# ---------------------------
# Treatment variables
# ---------------------------
hsls <- hsls %>%
  mutate(
    # adv_math: binary numeric (0 = not advanced, 1 = advanced)
    adv_math = case_when(
      x3thimath9 <= 4 ~ 1,
      x3thimath9 >= 5 ~ 0
    ),

    # math_3level: ordinal categorical (low → high)
    math_3level = ordered(
      case_when(
        x3thimath9 <= 3 ~ "1 remedial",
        x3thimath9 == 4 ~ "2 on-level",
        x3thimath9 >= 5 ~ "3 advanced"
      ),
      levels = c("1 remedial", "2 on-level", "3 advanced")
    ),

    # math_4level: ordinal categorical (low → high)
    math_4level = ordered(
      case_when(
        x3thimath9 <= 3 ~ "1 remedial",
        x3thimath9 == 4 ~ "2 on-level",
        x3thimath9 == 5 ~ "3 advanced",
        x3thimath9 >= 6 ~ "4 double advanced"
      ),
      levels = c("1 remedial", "2 on-level", 
                 "3 advanced", "4 double advanced")
    )
  )

# ---------------------------
# Controls
# ---------------------------
hsls <- hsls %>%
  mutate(
    # urban: binary numeric
    urban = ifelse(x1locale == 1, 1, 0),

    # math_score: continuous numeric
    math_score = ifelse(x1txmquint < 0, NA, x1txmtscor),
    math_quintile = x1txmquint,

    # public: binary numeric (1 = public, 0 = private)
    public = case_when(
      x1control == 1 ~ 1,
      x1control == 2 ~ 0
    )
  )

# ---------------------------
# Demographics
# ---------------------------
hsls <- hsls %>%
  mutate(
    # first_gen: binary numeric
    first_gen = ifelse(x1paredu <= 3, 1,0),
    par_ba = ifelse(x1paredu >= 4, 1, 0),

    # female: binary numeric (1 = female)
    female = ifelse(x1sex == 2, 1, 0),

    # frpl: binary numeric
    frpl = ifelse(x1poverty185 == 0, 0, 1),

    # pov_level: ordinal numeric (1 = deepest poverty → 4 = no poverty)
    pov_level = ordered(
      case_when(
        x1poverty185 == 0  ~ 4,
        x1poverty185 == 1 & x1poverty130 == 0 ~ 3,
        x1poverty130 == 1 & x1poverty == 0 ~ 2,
        x1poverty == 1 ~ 1
      ),
      levels = c(1, 2, 3, 4)
    ),

    # first_lang: nominal categorical
    first_lang = factor(
      case_when(
        x1duallang == 1 ~ "English only",
        x1duallang == 2 ~ "Non-English only",
        x1duallang == 3 ~ "English and another equally"
      )
    ),

    # esl_program: binary numeric
    esl_program = case_when(
      p1ellever == 1 ~ "1 yes",
      p1ellever == 2 ~ "0 no"
    ),

    # stud_race: nominal categorical
    stud_race = factor(
      case_when(
        x1race %in% c(1, 7) ~ "Indigenous",
        x1race == 2        ~ "Asian",
        x1race == 3        ~ "Black",
        x1race %in% c(4,5) ~ "Latinx",
        x1race == 6        ~ "Multiple",
        x1race == 8        ~ "White"
      )
    ),

    # par_race: nominal categorical
    par_race = factor(
      case_when(
        x1par1race == 2        ~ "Asian",
        x1par1race == 3        ~ "Black",
        x1par1race %in% c(4,5) ~ "Latinx",
        x1par1race == 6        ~ "Multiple",
        x1par1race == 8        ~ "White",
        x1par1race == 9        ~ "Indigenous"
      )
    )
  )

# ---------------------------
# Parent involvement
# ---------------------------
hsls <- hsls %>%
  mutate(

    # par_sch_inv: count variable (numeric, Poisson-type)
    par_sch_inv = p1ptomtg +
                  p1schmtg +
                  p1schevent +
                  p1volunteer +
                  p1fundraise +
                  p1ptconfer
  )

# ---------------------------
# Parent expectations
# ---------------------------
hsls <- hsls %>%
  mutate(
    exp_ba = ifelse(p1eduexpect >= 6, 1, 0),
    # par_exp: ordinal numeric (low → high expectations)
    par_exp = ordered(
      case_when(
        p1eduexpect == 11            ~ 1, # IDK
        p1eduexpect %in% c(1,2,3,4)  ~ 2, # high school to Associate
        p1eduexpect %in% c(5,6)      ~ 3, # start/complete BA
        p1eduexpect %in% c(7,8)      ~ 4, #start/complete MA
        p1eduexpect %in% c(9,10)     ~ 5 # sart/complete phd/professional
      ),
      levels = c(1, 2, 3, 4, 5)
    )
  )

# check poverty groups
##hsls %>% dplyr::select(pov_level, starts_with("x1pov")) %>% tbl_summary(by = pov_level)
```
# Save Clean Dataset

```{r save}
write_dta(hsls, "gcq_hsls_subset.dta")
```


# Analysis & Model Building on Unimputed Data

## Descriptive Statistics

```{r}
df_summ <- hsls %>% dplyr::select(par_says_honors, math_3level, 
                        exp_ba, par_ba, par_race, par_sch_inv,
                        first_lang, frpl, stud_race,
                        math_score, public, urban)

df_summ %>% gtsummary::tbl_summary(by = stud_race) %>% as_flex_table()
```

## Weight Data
```{r}
library(survey)

df_svy <- svydesign(
  data = hsls,
  ids = ~stu_id,      # Use the tilde (~) for formulas
  weights = ~w1parent # Use 'weights' instead of 'pweights'
)
df_brr <- svrepdesign(
  weights = ~w1parent, 
  repweights = "w1parent[0-9]+",
  type = "BRR", 
  data = hsls,
  combined.weights = TRUE
)
```

## Run Data for All 3 Math Levels
```{r}
glm_all3 <-glm(data = hsls,
           p1honors ~ math_3level + 
           par_exp + par_race + par_sch_inv +
           first_gen + first_lang + frpl +
           math_score + public + urban,
           family = "binomial")
jtools::summ(glm_all3, exp = TRUE)
glm_all3 %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% add_glance_table() %>% as_flex_table()
```

## Run Data for All 4 Math Levels
```{r}
glm_all4 <-glm(data = hsls,
           p1honors ~ math_4level + 
           par_exp + par_race + par_sch_inv +
           first_gen + first_lang + frpl +
           math_score + public + urban,
           family = "binomial")
jtools::summ(glm_all4, exp = TRUE)


glm_all4 %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% add_glance_table() %>% as_flex_table()
```

## Run Data for All 4 Math Levels with quintile instead of score
```{r}
glm_all4_quint <-glm(data = hsls,
           p1honors ~ math_4level + 
           par_exp + par_race + par_sch_inv +
           first_gen + first_lang + frpl +
           ordered(math_quintile) + public + urban,
           family = "binomial")
jtools::summ(glm_all4_quint, exp = TRUE)


glm_all4_quint %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% add_glance_table() %>% as_flex_table()
```

## Weighted Analysis for All: Panel Weights
```{r}
glm_all_panel <-       svyglm(design = df_svy,
                        p1honors ~ math_3level + 
                        par_exp + par_race + par_sch_inv +
                        first_gen + first_lang + frpl +
                        math_score + public + urban,
                        family = "binomial")
glm_all_panel %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% as_flex_table()
```

## Weighted Analysis for All: BRR weights
```{r}

glm_all_brr <-       svyglm(design = df_brr,
                        p1honors ~ math_3level + 
                        par_exp + par_race + par_sch_inv +
                        first_gen + first_lang + frpl +
                        math_score + public + urban,
                        family = "binomial")
glm_all_brr %>% tbl_regression(exp = TRUE) %>% add_significance_stars(hide_ci = FALSE, hide_se = FALSE, hide_p = FALSE) %>% as_flex_table()
```





